---
title: "ds202_lab4"
author: "Jonathan DeGardner"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(readxl)
library(tidyverse)
library(tidyr)

defense <- read_excel('C:\\Users\\Jonny DeGardner\\Desktop\\Cawlige\\S.S1\\DS 202\\ds202_lab4\\cyclonesFootball2019.xlsx', sheet='Defensive')

offense <- read_excel('C:\\Users\\Jonny DeGardner\\Desktop\\Cawlige\\S.S1\\DS 202\\ds202_lab4\\cyclonesFootball2019.xlsx', sheet='Offensive')

bio <- read_excel('C:\\Users\\Jonny DeGardner\\Desktop\\Cawlige\\S.S1\\DS 202\\ds202_lab4\\cyclonesFootball2019.xlsx', sheet='Biography')
```

## Part One: Cleaning Data
```{r}
defClean <- defense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Tackles_Solo = as.numeric(Tackles_Solo),
         Tackles_ASST = as.numeric(Tackles_ASST),
         Tackles_TFL = as.numeric(Tackles_TFL),
         Tackles_Sack = as.numeric(Tackles_Sack),
         Turnover_FF = as.numeric(Turnover_FF),
         Turnover_FR = as.numeric(Turnover_FR),
         Turnover_INT = as.numeric(Turnover_INT),
         Pass_QBH = as.numeric(Pass_QBH),
         Pass_PB = as.numeric(Pass_PB))
str(defClean)

offClean <- offense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Rushing_ATT = as.numeric(Rushing_ATT),
         Rushing_YDS = as.numeric(Rushing_YDS),
         Rushing_TD = as.numeric(Rushing_TD),
         Receiving_REC = as.numeric(Receiving_REC),
         Receiving_YDS = as.numeric(Receiving_YDS),
         Receiving_TD = as.numeric(Receiving_TD),
         `Passing_CMP-ATT` = as.numeric(`Passing_CMP-ATT`),
         Passing_YDS = as.numeric(Passing_YDS),
         Passing_TD = as.numeric(Passing_TD),
         Passing_INT = as.numeric(Passing_INT))
str(offClean)

bioClean <- bio %>%
  separate(Height, c('Feet', 'Inches'), '-', convert=TRUE) %>%
  mutate(Name = as.factor(Name),
         Position = as.factor(Position),
         Feet = as.numeric(Feet),
         Inches = as.numeric(Inches),
         Weight = as.numeric(Weight),
         Class = as.factor(Class),
         Hometown = as.factor(Hometown),
         Highschool = as.factor(Highschool))
str(bioClean)
```

## Part Two: Tidying
#### Question 1
```{r}
defClean2 <- defClean %>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to='stat')
```

#### Question 2
```{r}
ggplot(defClean2, aes(x=stat, y=value)) + geom_bar(stat='identity', width=.75, color='red') + theme(axis.text.x = element_text(angle=90))
```

The three least common defensive skills are: Turnover_FF (fumble forced), Turnover_FR (fumble recovery), and Turnover_INT (interception).

#### Question 3
```{r}
iowa <- defClean %>%
  filter(Opponent_Opponent=='Iowa') %>%
  ggplot(aes(x=Name, y=Tackles_Solo)) + geom_point() + theme(axis.text.x = element_text(angle=90))
iowa

nd <- defClean %>%
  filter(Opponent_Opponent=='Notre Dame') %>%
  ggplot(aes(x=Name, y=Tackles_Solo)) + geom_point() + theme(axis.text.x = element_text(angle=90))
nd
```

ISU had a better defense against Notre Dame, as there are more players with a larger amount of solo tackles in this game.

#### Question 4
```{r}
sep <- bioClean %>%
  separate(Hometown, c('City', 'State'), sep = ',')
head(sep)
```

#### Question 5
```{r}
sep %>%
  count(State)
```

ISU football players are from 19 different states. The table above shows how many players are from each state.

## Part Three: Joining Data Frame
#### Question 1
```{r}
def_state <- inner_join(x=defClean, y=sep)
def_state %>%
  count(State)

off_state <- inner_join(x=offClean, y=sep)
off_state %>%
  count(State)
```

The first table shows the count of defensive ISU football players from each state.
The second table shows the count of offensive ISU football players from each state.

#### Question 2
```{r}
p3q2 <- full_join(x=offClean, y=defClean)
purdy <- p3q2 %>%
  filter(Name == 'Purdy, Brock')

rest <- p3q2 %>%
  filter(Name != 'Purdy, Brock')

ggplot(purdy, aes(x=Opponent_Opponent, y=Rushing_TD)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))
ggplot(purdy, aes(x=Opponent_Opponent, y=Passing_TD)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))

ggplot(rest, aes(x=Opponent_Opponent, y=Rushing_TD)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))
ggplot(rest, aes(x=Opponent_Opponent, y=Receiving_YDS)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))
ggplot(rest, aes(x=Opponent_Opponent, y=Turnover_INT)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))
ggplot(rest, aes(x=Opponent_Opponent, y=Tackles_Sack)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))

```

#### Question 3
```{r}
defense2018 <- read_excel('C:\\Users\\Jonny DeGardner\\Desktop\\Cawlige\\S.S1\\DS 202\\ds202_lab4\\cyclonesFootball2018.xlsx', sheet='Defensive')
defClean2018 <- defense2018 %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Tackles_Solo = as.numeric(Tackles_Solo),
         Tackles_ASST = as.numeric(Tackles_ASST),
         Tackles_TFL = as.numeric(Tackles_TFL),
         Tackles_Sack = as.numeric(Tackles_Sack),
         Turnover_FF = as.numeric(Turnover_FF),
         Turnover_FR = as.numeric(Turnover_FR),
         Turnover_INT = as.numeric(Turnover_INT),
         Pass_QBH = as.numeric(Pass_QBH),
         Pass_PB = as.numeric(Pass_PB))

offense2018 <- read_excel('C:\\Users\\Jonny DeGardner\\Desktop\\Cawlige\\S.S1\\DS 202\\ds202_lab4\\cyclonesFootball2018.xlsx', sheet='Offensive')
offClean2018 <- offense2018 %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Rushing_ATT = as.numeric(Rushing_ATT),
         Rushing_YDS = as.numeric(Rushing_YDS),
         Rushing_TD = as.numeric(Rushing_TD),
         Receiving_REC = as.numeric(Receiving_REC),
         Receiving_YDS = as.numeric(Receiving_YDS),
         Receiving_TD = as.numeric(Receiving_TD),
         `Passing_CMP-ATT` = as.numeric(`Passing_CMP-ATT`),
         Passing_YDS = as.numeric(Passing_YDS),
         Passing_TD = as.numeric(Passing_TD),
         Passing_INT = as.numeric(Passing_INT))


defClean2018_2 <- defClean2018 %>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to='stat')
defClean2 <- defClean %>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to='stat')
offClean2018 <- offClean2018 %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to='stat')
offClean2 <- offClean %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to='stat')

ggplot(defClean2018_2, aes(x=Name, y=value)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))
ggplot(defClean2, aes(x=Name, y=value)) + geom_bar(stat='identity', width=.75, color='blue') + theme(axis.text.x = element_text(angle=90))
ggplot(offClean2018, aes(x=Name, y=value)) + geom_bar(stat='identity', width=.75, color='red') + theme(axis.text.x = element_text(angle=90))
ggplot(offClean2, aes(x=Name, y=value)) + geom_bar(stat='identity', width=.75, color='red') + theme(axis.text.x = element_text(angle=90))
```

